---
title: "BC LDB Beer Sales - Explore"
author: "`r Sys.getenv('USER')`"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    code_fold: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message=FALSE,
                      warning=FALSE,
                      fig.height=3.5,
                      fig.width=8)

library(tidyverse)
library(lubridate)
library(here)
library(PerformanceAnalytics)
library(plotly)
library(scales)
library(gridExtra)
library(DT)
library(RMariaDB)
theme_set(theme_bw())
```

## Intro 

Exploring data collected so far on LDB beer sales. Data collected in LMR-fetch-process.R.

## Import data from MySQL

```{r}
## get from database
## from file in .gitignore
source('credo.R')
con <- dbConnect(RMariaDB::MariaDB(), user='root', password=mypwd, dbname='bcbg')
#dbGetQuery(con, "SELECT * FROM tblLDB_beer_sales;")
## get list of tables (if needed to confirm)
#dbGetQuery(con, "SHOW TABLES;")
## import data into dataframe
lmr_data <- dbGetQuery(con, "SELECT * FROM tblLDB_beer_sales;")
str(lmr_data)
```

## Summarize

```{r}
summary(lmr_data)
cat('Date coverage: \n')
table(lmr_data$qtr, lmr_data$fyr)
#table(lmr_data$category, lmr_data$subcategory)
cat('Category > Subcategory: \n')
lmr_data %>% group_by(category, subcategory) %>% summarize(sales=sum(netsales),
                                                           min_date=ymd(min(end_qtr_dt)),
                                                           max_date=ymd(max(end_qtr_dt)))
## trying to identify metric based on excluded non-metric dimensions
cat('Metrics used \n')
metric <- lmr_data %>% select(-idbeersales, 
                    -category,
                    -subcategory,
                    -period,
                    -qtr,
                    -fyr,
                    -end_dt,
                    -end_qtr_dt,
                    -cyr) 
names(metric)
```

## Trends

```{r}
ch_title <- "Total Sales by Qtr"
lmr_trend <- lmr_data %>% group_by(end_qtr_dt) %>% summarize(sales=sum(netsales)/1000000)
lmr_trend %>% ggplot(aes(x=end_qtr_dt, y=sales))+geom_line()+
  scale_y_continuous(labels=comma, limits=c(0,max(lmr_trend$sales)))+
  geom_smooth(method=lm)+
  labs(title=ch_title, x="", y="$ sales (millions)")
```

Observation:

* slight downward trend, with lots of consistent seasonality

## Sales by Category

```{r}
## breakdown by main categories
## sales numbers
ch_title <- "Sales by Category"
lmr_data %>% ggplot(aes(x=period, y=netsales/1000000, fill=category))+geom_col()+
  scale_y_continuous(labels=comma, expand=expansion(mult=c(0,0.05)))+
  labs(title=ch_title, x="", y="$ sales (millions)")+
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1),
        axis.ticks.x=element_blank())

```

```{r}
## % of total by category
ch_title <- "% Sales by Category"
lmr_data %>% ggplot(aes(x=period, y=netsales, fill=category))+geom_col(position='fill')+
  scale_y_continuous(labels=percent_format(), expand=expansion(mult=c(0,0)))+
  labs(title=ch_title, x="", y="$ sales (millions)")+
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1),
        axis.ticks.x=element_blank())
  
## observation: domestic - bc beer is dominant and stable ~75%
```

Observation:

* domestic bc beer is dominant and stable

## Zoom in on Domestic - BC Beer

```{r}
## zoom in on domestic - bc beer
ch_title <- "Domestic BC Beer Sales by Subcategory"
lmr_data %>% filter(category=='Domestic - BC Beer') %>% 
  ggplot(aes(x=period, y=netsales/1000000, fill=subcategory))+geom_col()+
  scale_y_continuous(labels=comma, expand=expansion(c(0,0.05)))+
  labs(title=ch_title, x="", y="$ sales (millions)")+
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1),
        axis.ticks.x=element_blank())

## % of total by category
ch_title <- "% Domestic BC Beer Sales by Subcategory"
lmr_data %>% filter(category=='Domestic - BC Beer') %>% 
  ggplot(aes(x=period, y=netsales/1000000, fill=subcategory))+geom_col(position='fill')+
  scale_y_continuous(labels=percent_format(), expand=expansion(c(0,0)))+
  labs(title=ch_title, x="", y="$ sales (millions)")+
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1),
        axis.ticks.x=element_blank())

## observation: also stable, dominated by commercial at ~60%; micro brew ~15%
```

Observation:

* dominated by commercial at ~60%, but shrinking
* regional biggest gainer (could depend on definitions, movements from 'micro' to 'regional')
* micro brew ~15%

```{r}
ch_total <- "% Share of Domestic BC beer subcategories"
lmr_pc <- lmr_data %>% filter(category=='Domestic - BC Beer') %>% 
  group_by(end_qtr_dt, period, subcategory) %>%
  summarize(sales=sum(netsales)) %>%
  mutate(
    pc_total=sales/sum(sales)
  )
lmr_pc %>% ggplot(aes(x=end_qtr_dt, y=pc_total, color=subcategory))+geom_line()+
  scale_y_continuous(labels=percent_format(), expand=expansion(c(0,0.05)))+
  labs(title=ch_title, x="", y="% Domestic BC Beer Sales")
```

```{r}
ch_total <- "% Share of Domestic BC beer subcategories"

lmr_pc %>% ggplot(aes(x=period, y=pc_total, fill=subcategory))+geom_col(position=position_dodge())+
  scale_y_continuous(labels=percent_format(), expand=expansion(c(0,0.05)))+
  labs(title=ch_title, x="", y="% Domestic BC Beer Sales")+
  theme(axis.text.x=element_text(angle=45, hjust=1, vjust=1),
        axis.ticks.x=element_blank())
```

